---
title: "131-Homework2"
output: html_document
author: 'Caleb Mazariegos'
date: '2022-04-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
abalone <- read.csv('/Users/calebmazariegos/Desktop/homework-2/data/abalone.csv')
library("tidyverse")
library("tidymodels")
```

Question 1:
Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no age variable in the data set. Add age to the data set.

Assess and describe the distribution of age.
```{r}
abalone <- abalone %>%
  mutate(age = rings + 1.5)

ggplot(abalone, aes(age)) + geom_bar()
```

Age has a normal distribution that is skewed right. 


Question 2:
Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

Remember that you’ll need to set a seed at the beginning of the document to reproduce your results.

```{r}
set.seed(3435)

abalone_split <- initial_split(abalone, prop = 0.80, strata = age)

abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```


Question 3
Using the training data, create a recipe predicting the outcome variable, age, with all other predictor variables. Note that you should not include rings to predict age. Explain why you shouldn’t use rings to predict age.

```{r}
abalone_recipe <- recipe(age ~ ., data = abalone_train %>% select(- rings)) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~starts_with("type"): shucked_weight)%>%
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms= ~ shucked_weight:shell_weight) %>%
step_normalize() %>% step_center(all_predictors()) %>%
step_scale()
abalone_recipe
```

We should not use rings to predict age because since age is 1.5 + rings, the prediction will be perfect.

Question 4:
Create and store a linear regression object using the "lm" engine.

```{r}
lm_model <- linear_reg() %>%
  set_engine("lm")
```
Question 5:
Now:

1. set up an empty workflow,
2. add the model you created in Question 4, and
3. add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(abalone_recipe)
```

Question 6:
Use your fit() object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_train)
lm_fit

lm_fit %>%
  extract_fit_parsnip() %>%
  tidy()

random_data <- data.frame(type = "F", longest_shell = 0.50,
diameter = 0.10, height = 0.30, whole_weight = 4,
shucked_weight = 1, viscera_weight = 2, shell_weight = 1,
stringsAsFactors = TRUE)

prediction <- predict(lm_fit, new_data = random_data)
prediction
```


Question 7:
Now you want to assess your model’s performance. To do this, use the yardstick package:

1. Create a metric set that includes R2, RMSE (root mean squared error), and MAE (mean absolute error).

2. Use predict() and bind_cols() to create a tibble of your model’s predicted values from the training data along with the actual observed ages (these are needed to assess your model’s performance).

3. Finally, apply your metric set to the tibble, report the results, and interpret the R2 value.


```{r}
library("yardstick")

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res %>%
  head()

abalone_train_res <- bind_cols(abalone_train_res, new_data = abalone_train %>% select(age))
abalone_train_res %>%
  head()

rmse(abalone_train_res, truth = age, estimate = .pred)

abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, estimate = .pred)
```



